---
title: "STSM_report_01"
author: "Marco Baldo"
date: "2026-06-20"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # hide code
  message = FALSE, # hide messages
  warning = FALSE  # hide warnings
)
```

# Data Initial Rendering & Saving

path: C:\P\DMP_CROSS_CASCADE\03_rawdata\02_process_storage

- Step 01: Load and prepare the Field Inventory Data 2015_2019 **FID_2015_2019** and 2022 **FID_2022** complete with species names converted, dead and alive trees and their condition = cond (if snag, broken, dry etc.).

1a. split the dead and alive trees in 0 - 1 dead/alive 2015_2019 all in the same data frame **FID_2015_2019_D**

1b. split the dead and alive trees in 0 - 1 dead/alive 2022 into the same data frame **FID_2022_D**

1c. Create six separate df 2015 - 2019 & alive - dead
**FID_2015** all trees but 2015
**FID_2019** all trees but 2019
**FID_2015_alive** only alive 2015
**FID_2015_dead** only dead 2015
**FID_2019_alive**
**FID_2019_dead**

1d. 2022 alive / dead (same logic as above)
**FID_2022_alive**
**FID_2022_dead**

1e. Check the whole n. of tree presence in 2015-2019 and the n. removal - regrowth (tree different between years).

1f. Inspect and create a df of trees (removal - regrowth) comparable between them. **problem_cases**

1g. Create a data frame to make estensive comparison analysis: time-series modeling - mixed models - growth/mortality transitions analysis. **FID_2015_2019_long**

1h. FID 2015 vs 2019 differences | FID 2015 vs 2022 differences | FID 2019 vs 2022 differences
    The 2022 plots are always present in 2015 & 2019 dataset
    

- Step 02: Save all the data frame separetly and in .csv & .rds (rds is the format to get injected in R directly as dataframe in the environment)


- Step 03: Extract the ALS based data from the GIS shape file - Table of attribute. NB: here the volume isn't present. Only Important information column structure for init table in iLand: x y species dbh height **ALS_2015**

3a. Check the species names and figure it out the corrispondence of not consisten names

3b. Make a consistent species name, including dead trees where needed, using the reference of the GEE article based species names: <https://gee-bialowiezamaps.projects.earthengine.app/view/bialowiezamapviewer>

reference: Lisiewicz, M., Kamińska, A., Kraszewski, B., Kuberski, Ł., Pilch, K., Stereńczak, K. 2025. Comprehensive mapping of individual living and dead tree species using leaf-on and leaf-off ALS and CIR data in a complex temperate forest. Forestry: An International Journal of Forest Research. <https://doi.org/10.1093/forestry/cpaf007>

3c. # fix ALS plot IDs to match FID format (KS001). contingency variables column names between FID and ALS using the FID: **ALS_2015**

3d. Check and SAVE MATCHING CONTROL PLOTS FILE between FID 2015 and ALS 2015: **matching_plots.txt**, **FID_not_in_ALS.txt**, **ALS_not_in_FID.txt**.

3e. Let's filter the data frame for FID 2015 - FID 2019 - ALS 2015 in the way to keep only **matching plots 668**

**FID_match** 2015 only
**ALS_match** 2015

3f. Define common plot set across ALL datasets ALS 2015 - FID 2015 and 2019 **valid_plots**

3g. filter ALL datasets to same plots and create the df: **FID_2015_clean**, **FID_2019_clean**, **ALS_2015_clean**. 

**THIS ARE THE IMPORTANT DATAFRAMES**

3h. Let's create different data frame for alive and dead trees as in FID and for ALS 2015 same plots
**ALS_clean_alive**
**ALS_clean_dead**
**FID_2015_clean_alive**
**FIS_2015_clean_dead**
**FID_2019_clean_alive**
**FIS_2019_clean_dead**

3i. Save in csv and rds all the clean data set and ALS based dataset including the full

- Step 4: Global Test

4a. Test if is true that the 2015 2019 and ALS 2015 dataset have the same plot ids. Make a random check o then number of trees in some selected plots.

4b. Import the FID plots information table with age and dominant species and edit the df according to species names and number of matching plots **FID_2015_2019_Info_clean**

4c. TEST OF PLOTS NUMBERS AND TREE COUNTS **summary_plot**, **plot_count_check**

```{r FID data, echo=FALSE, message=FALSE, warning=FALSE}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
